Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-04-17 apj.tex
more changes from Greg's comments

Differences between the model-predicted versus
\color{orange}
predicted
\color{black}
 supernova colors are shown in Figure~\ref{residual:fig}.
There is no evidence of a catastrophic fit nor of extreme
outlying data.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/residual.pdf} 
            \caption{Differences between predicted versus predicted colors versus the \color{orange}
predicted
\color{black} color.  For each color, the ideogram of the color differences is
            shown on the right.
            \label{residual:fig}}
\end{figure}

@@

\color{orange}
For brevity, we do not present here the analysis of simulated data used to validate and bias quantify bias in our results.
The simulation tests in \S\ref{model2:sec} are representative for both Model~1 of this Section and Model~2 of the subsequent section.
\color{black}

@@

We compare our model and its color-coefficient parameters $\gamma^0$ and $\gamma^1$ with the expectations of  dust models.
The 2-parameter linear model
\begin{equation}
A_X = a(X)  A_V + b(X) E(B-V)
\label{f99:eqn}
\end{equation}
approximately describes dust models.
For the case of
$R^F=2.5$ and $A^F_V=0.1$,
and the measurement of the SALT2
\citet{2007A&A...466...11G} $s=1$, $x_1=0$ SN~Ia template at $B$-band peak, the \citetalias{1999PASP..111...63F} model
gives
$a(U,B,V,R,I)=\{0.96,   1.00,   1.00,   0.97,   0.77\}$ and $b(U,B,V,R,I)=\{  1.77,   0.98,   0.12,  -0.50,  -0.53\}$.
The $F$ superscript is added to distinguish \citetalias{1999PASP..111...63F} parameters.
Over the ranges of
 $R^F$ and $A^F_V$,
and the wavelengths under consideration,  the values of the elements of $a$ and $b$ change but only by $<5$\%.

@@

This result indicates that
the allowed variation in $UBVRI$ allowed by the \citetalias{1999PASP..111...63F} model and our best-fit model are closely aligned.
That is to say, with no prior assumptions of dust extinction behavior or the distribution of $A_V$, the supernova data themselves exhibit
2-dimensional color variations that are closely aligned with the 2-dimensional color variations predicted by  \citetalias{1999PASP..111...63F} dust.

@@

When fixing the origin of the $\gamma$'s to zero, the are found to be coplanar with $a$ and $b$.
However the linear model does not specify zero dust so the origin of gamma is unspecified.
The general result is that the magnitude planes of the linear and dust models are parallel.
Color changes are consistent between the two models but there is no specification of the relation between $k^0$ and $k^1$ with $A_V$ and $E(B-V)$.

A prior that $k_0$ and  $k_1$ be uncorrelated was enforced to break degeneracies in our linear models, and determines the choice of directions of
the vectors $\gamma^0$ and $\gamma^1$.  Suppose that that the data were described by a dust model.  A \citetalias{1999PASP..111...63F}  dust model 
can be written as 
%To see how this prior would operate on dust models, consider that themodel in Eqn.~\ref{f99:eqn} can
%be expressed in terms of new basis vectors parametrized by $\kappa$:
\begin{equation}
A_X =  (1-\kappa_1 \kappa_2)^{-1} [(A^F_V - \kappa_2 E^F(B-V))\left(a(X)+\kappa_1 b(X) \right) +  (-\kappa_1 A^F_V + E^F(B-V)) (\kappa_2 a(X) + b(X))],
\end{equation}
with the parameters $\kappa_1$ and $\kappa_2$ providing freedom in the directions of the two basis vectors.
The vector $a(X)+\kappa_1 b(X)$ represents extinction that can occur with fixed $R^F_0=1/\kappa_1$ and unconstrained $\kappa_2$.
%While a similar interpretation could be made of $\kappa_2 a(X) + b(X)$, note that to first order the extinction for $\Delta R$
%perturbations from a fixed $R^F_0$ is
%\begin{equation}
%A_X = A_V \left(a(X) + \left(\frac{1}{R^F_0} - \frac{\Delta R^F}{R^F_0} \right) b(X) \right).
%\end{equation}
%Then $\kappa^{-1}_2$ can this be associated with $\frac{1}{R^F_0} - \frac{\Delta R^F}{R^F_0}$.

@@

\color{orange}
As a check of the above interpretation, we generate ten simulated datasets with the same numbers and uncertainties as our data, based on a
SALT2 $x_1=0$, $c=0$ template and  \citetalias{1999PASP..111...63F} dust.  The supernovae
are assigned $E^F(B-V)$ parameters equal to the $B-V$ color excesses determined in our sample, and $R^F$ values drawn from a log-normal
distribution with a mean of $\ln{2.5}$ and 0.05 standard deviation.  The $\alpha$, $\beta$, and $\eta$ parameters are set to zero, so that dust and
measurement uncertainty are the only sources of variation in magnitudes.  The input 2.5 is centered
within the 68\% credible interval for $\kappa^{-1}_1$ of 2.38--2.69 from the results of the stacked 10 simulations.

@@

Figure~\ref{k0_med:fig} also shows the distribution of $\gamma^1_V k_1$, which can be attributed to 
extinction corrections due to deviations away from the canonical $\kappa_1$ value.   The distribution has a long
negative tail and is peaked at its positive edge, and by construction has a mean of zero.
\color{orange}
Note that the extreme outlier has large uncertainties, and is a manifestation of the unconstrained degree of freedom in this parameter set.
\color{black}
The implication is that while the bulk of supernovae have positive values associated with $R>1/\kappa_1$, there is a tail
of negative values associated with $R<1/\kappa_1$.  The plot of   $\gamma^0_V k_0$ versus $\gamma^1_V k_1$  shown in Figure~\ref{kk:fig} shows that
the positive edge in $\gamma^1_V k_1$ persists over the full range of  $\gamma^0_V k_0$. On the other hand, lower values of $\gamma^1_V k_1$
occur at larger values of  $\gamma^0_V k_0$.
These findings are consistent with previous results:
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F<3$ encountered by the SN~Ia population.
\citet{2011ApJ...729...55F} and the above authors
find that supernovae with large extinction 
preferentially exhibit low $R^F$.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output11/gamma0_med.pdf}
   \includegraphics[width=2.8in]{output11/gamma1_med.pdf}
      \caption{
      \color{orange} Normalized ideogram and histogram 
      \color{black}
      of
median values of Left: $\gamma^0_V k_0$, which is associated up to an offset with the $A^F_V$ of an $R^F \sim 1/\kappa_1$ dust model; Right: $\gamma^1_V k_1$,
which  is associated with extinction corrections due to deviations away from the canonical $\kappa_1$ value.
   \label{k0_med:fig}}
\end{figure}

@@

\subsection{SN 2014J in the Context of Our Model}
\label{sn2014j:sec}
SN~2014J   is one of several SNe~Ia that exhibits colors that imply a low $R_V<2.0$ \citep{2014ApJ...788L..21A, 2014MNRAS.443.2887F, 
2014arXiv1411.3332J,
2014ApJ...795L...4K, 2015ApJ...805...74B}.
The colors of this supernova can be studied in comparison to SN~2011fe, an object with similar
spectral evolution that 
suffers very low galactic and host reddening
\citep[this technique has been used in][]{2006MNRAS.369.1880E,2007AJ....133...58K,2008MNRAS.384..107E,2010AJ....139..120F, 2014ApJ...788L..21A,
2017arXiv170101422H}.
The overall higher photospheric velocities of
SN~2014J are unimportant  based on the findings of  \S\ref{results:sec}.

Data of SN~2014J are  analyzed using our model and using the parameter values found with the SNfactory sample.
The data for the color excesses  in $UBRi$  relative to $V$ at peak brightness  are taken from \citet{2014ApJ...788L..21A},
following their prescription of averaging measurements within 5-days of peak $B$ magnitude.
Their values 
$E_o(U-V) =   2.23 \pm   0.03$,
$E_o(B-V) =   1.28 \pm   0.04$,
$E_o(R-V) =  -0.47 \pm   0.03$,
$E_o(i-V) =  -0.92 \pm   0.03$
are plotted in Figure~\ref{sn2014j:fig}.
The central wavelengths of the filters are 3660, 4280, 5350, 6310, and 8100 \AA\ respectively, in contrast to the central
wavelengths 
3640, 4390, 5290, 6370, 7680 \AA\ of our synthetic bands.  Despite the imperfect correspondence our synthetic passbands
and those used for the observations,
these data are fit to our model
\begin{equation}
E_o(X-V) =  \left(\frac{\gamma^0_X}{\gamma^0_V}-1\right)e_0 +  \left(\frac{\gamma^1_X}{\gamma^1_V}-1\right)e_1,
\end{equation}
where we use the median values of the $\gamma$- and $\delta$-terms from Table~\ref{global:tab} and the fit
parameters $e_i $.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/sn2014j.pdf} 
   \caption{Top: Measured color excess for SN~2014J and the best-fit predictions from this article.   Bottom: Residual color excess
   between the data and model.
   This can be compared with the best-fit results of
   \citet{2014ApJ...788L..21A} shown in their Figure~3.
   \label{sn2014j:fig}}
\end{figure}

The best-fit parameters for SN~2014J are 
%----
$x_0= 2.60$, $ x_1=-1.89$ with covariance
%----
\begin{equation}
\begin{pmatrix}
\begin{array}{rr}
0.050 & 0.058 \\
0.058 & 0.315
\end{array}
\end{pmatrix}.
\end{equation}
The predicted values of observed $E_o(B-V)$ from the fit are shown in Figure~\ref{sn2014j:fig}, where they are found to
overlap with the data within the error bars.   
Our results can be compared with those of  \citet{2014ApJ...788L..21A} (Figure 3), who 
use
UV through NIR data to fit \citet{1999PASP..111...63F} dust with $R_V^F=1.4$ and $E(B-V)=1.37$.
Our model predictions provide a significantly better match to the data
despite the only approximate correspondence of the model and observer passbands.

@@

\section{Supernova Model II: Two Extrinsic Parameters, One Intrinsic Parameter}
\label{model2:sec}
In \S\ref{model:sec} we show that the SNfactory data analyzed with an agnostic model with
two extrinsic parameters produce
magnitudes consistent
with two-parameter dust models that have been established for stars in the Milky Way, LMC, and SMC.
An extension to our model is now added to explore the possibility of
a possible additional
color parameter that could have its origins in supernova astrophysics.
The extended model is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_U+\alpha_U EW_{Ca} + \beta_U EW_{Si} + \eta_U \lambda_{Si} + \delta_U D\\
c_B+\alpha_B EW_{Ca} + \beta_B EW_{Si} + \eta_B \lambda_{Si} + \delta_B D \\
c_V+\alpha_V EW_{Ca} + \beta_V EW_{Si} + \eta_V \lambda_{Si} + \delta_V D\\
c_R+\alpha_R EW_{Ca} + \beta_R EW_{Si} + \eta_R \lambda_{Si} + \delta_R D\\
c_I+\alpha_I EW_{Ca} + \beta_I EW_{Si}+ \eta_I \lambda_{Si} + \delta_I D
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{U} k_0 +\gamma^1_{B} k_1 \\B +\gamma^0_{B} k_0 +\gamma^1_{B} k_1 \\
V+\gamma^0_{V} k_0+\gamma^1_{V} k_1\\R+\gamma^0_{R} k_0 + \gamma^1_{R} k_1\\I+\gamma^0_{I} k_0+\gamma^1_{I} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
In all aspects it is identical to the previous model except for the addition of the term $\delta D$,
which affects the mean
absolute magnitudes.  Thus
each supernova is given an intrinsic $D$ parameter that is linearly related to
absolute magnitudes through global coefficients $\delta$.  As with $k_0$ and $k_1$,  $D$ is a latent supernova parameter whose existence is inferred from
its effects on broadband magnitudes.

@@

To constrain the new degrees of freedom and degeneracies inherent in the model we impose additional constraints. We impose
\begin{equation}
\langle D\rangle=0,\,  \delta \cdot \gamma^0=0,\, \delta \cdot \gamma^1=0.
\end{equation}
The first of these constraints specifies the definition of zero color.   The latter two confine possible magnitude
variations to be those that cannot be attributed to 
the $\gamma k$ terms (i.e.\ dust); this condition breaks the degeneracies between $\delta D$  
and $\gamma k$. 
Unlike for the $\gamma$'s, there is no condition on the sign of $\delta_U$,
so that the posterior has degenerate solutions with sign flips of $\delta$ and $D$.
We apply the condition $\delta_I>0$ to break this degeneracy,
and assume a flat prior for the three linear components of the subspace perpendicular to the $\gamma$ plane.
In the running of the Monte Carlo fit, the initial conditions for
$\gamma^0$ and $\gamma^1$ were set to be in the neighborhood of the best fit of \S\ref{model:sec}
to maintain their correspondence between the two models.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output23/ebv.pdf}
   \includegraphics[width=2.8in]{output23/ebv_delta.pdf}
      \caption{Left: Ideograms of the external $E(B-V)$ and
   internal $E_\delta(B-V) = (\delta_B-\delta_V)D$  contributions to color excess  for the supernovae in our sample.
   Right: Ideogram of $E_\delta(B-V)$ on an enlarged scale.
   \label{ebv:fig}}
\end{figure}
The significance of the $\delta$ parameters is shown pictorially in
Figure~\ref{deltacorner:fig}, which shows the $\delta$ posteriors.
Recall that fixing $\delta_X=0$ reverts this model into the two extrinsic-parameter model of \S\ref{model:sec}.
Assigning the Monte Carlo links to bins of width 0.5 in each $\delta$ parameter, the maximum number of links in a single bin is 303, the number of bins
with at least one element is 1987, and the bin that contains $\delta=0$ for all bands has zero elements; we thus claim a detection
of a third color-parameter at $> (1-5\times 10^{-4})$.
Figure~\ref{deltaratio:fig} shows  values of $\delta_X/\delta_I-1$ that  have significant non-zero values.
While the signal of the internal parameter is strong within the context of this model, ``model testing'' (comparing the
performance  between this model
and that of \S\ref{model:sec}), is challenging for Bayesian Monte Carlo analysis.  Evidence calculations are planned
for the next generation of STAN and the development of robust algorithms is a subject of theoretical work.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
      \includegraphics[width=5.2in]{output23/deltaratio.pdf}
   \caption{Significance of the non-zero values of $\delta$ as demonstrated by the best-fit values of $\delta_X/\delta_I-1$ for $\delta$ for Model~II.
   \label{deltaratio:fig}}
\end{figure}

@@

\color{black}
The analysis is validated using simulated datasets with and without third color parameters: the input parameters are recovered within  uncertainties.
We present here the results of one of the validation tests.  The input values of the supernova parameters are the median values from the analysis of real data.
Ten realizations of simulated data are generated based on the data covariances, and then analyzed separately.   The posteriors of the ten realizations
are aggregated and shown in Figures~\ref{simglobal1:fig}--\ref{simglobal5:fig}.  The input parameters are recovered with negligible bias
relative to the statistical uncertainties.  There is a tendency for the slope parameters to be slightly underestimated, which is an expected consequence of 
the uncertainties in the predictor variable, which in turn lead to 
regression dilution.  Our analysis is susceptible to this bias because we do consciously do not include distributions
of supernova parameters as part of our model.

@@

\section{Discussion and Summary}

>>>>

\color{orange}
Observed versus deduced colors for four selected colors, and the histogram of their differences,
\color{black}
are shown in Figure~\ref{residual:fig}.
There is no evidence of a catastrophic fit nor of extreme
outlying data.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/residual.pdf} 
            \caption{
            \color{orange}
            Observed versus deduced colors for four selected colors.
On the right for each color is plotted the histogram of the difference between the observations and deductions.
\color{black}
            \label{residual:fig}}
\end{figure}

@@

\color{orange}
For brevity, we do not present here the analysis of simulated data used to validate and quantify bias in our results.
The simulation tests in \S\ref{model2:sec} are representative for both Model~1 of this Section and Model~2 of the subsequent section.
\color{black}

@@

We compare our model and its color-coefficient parameters $\gamma^0$ and $\gamma^1$ with the expectations of  dust models.
The 2-parameter linear model
\begin{equation}
A_X = a(X)  A_V + b(X) E(B-V)
\label{f99:eqn}
\end{equation}
approximately describes dust models \citep[e.g][]{1989ApJ...345..245C}.
For the case of
$R^F=2.5$ and $A^F_V=0.1$,
and the measurement of the SALT2
\citet{2007A&A...466...11G} $s=1$, $x_1=0$ SN~Ia template at $B$-band peak, the \citetalias{1999PASP..111...63F} model
gives
$a(U,B,V,R,I)=\{0.96,   1.00,   1.00,   0.97,   0.77\}$ and $b(U,B,V,R,I)=\{  1.77,   0.98,   0.12,  -0.50,  -0.53\}$.
The $F$ superscript is added to distinguish parameters of the \citetalias{1999PASP..111...63F} model.
Over the ranges of
 $R^F$ and $A^F_V$,
and the wavelengths under consideration,  the values of the elements of $a$ and $b$ change but only by $<5$\%.

@@

This result indicates that
the allowed variation in $UBVRI$ allowed by the \citetalias{1999PASP..111...63F} model and our best-fit model are closely aligned.
That is to say, with no prior assumptions of dust extinction behavior or the distribution of $A_V$, the supernova data themselves exhibit
2-dimensional color variations that are closely aligned with the 2-dimensional color variations predicted by the \citetalias{1999PASP..111...63F} dust model.

@@

When fixing the origin of the $\gamma$'s to zero, the $\gamma$ vectors are found to be coplanar with $a$ and $b$.
However the linear model does not specify zero dust so the origin of gamma is unspecified.
The general result is that the magnitude planes of the linear and dust models are parallel.
Color changes are consistent between the two models but there is no specification of the relation between $k^0$ and $k^1$ with $A_V$ and $E(B-V)$.

A prior that $k_0$ and  $k_1$ be uncorrelated was enforced to break degeneracies in our linear models, and determines the choice of directions of
the vectors $\gamma^0$ and $\gamma^1$.  Suppose that the data were described by a dust model.  Then a \citetalias{1999PASP..111...63F}  dust model 
can be written as 
%To see how this prior would operate on dust models, consider that themodel in Eqn.~\ref{f99:eqn} can
%be expressed in terms of new basis vectors parametrized by $\kappa$:
\begin{equation}
A_X =  (1-\kappa_1 \kappa_2)^{-1} [(A^F_V - \kappa_2 E^F(B-V))\left(a(X)+\kappa_1 b(X) \right) +  (-\kappa_1 A^F_V + E^F(B-V)) (\kappa_2 a(X) + b(X))],
\end{equation}
with the parameters $\kappa_1$ and $\kappa_2$ providing freedom in the directions of the two basis vectors.
The vector $a(X)+\kappa_1 b(X)$ represents extinction that can occur with fixed $R^F_0=1/\kappa_1$ and unconstrained $\kappa_2$.
%While a similar interpretation could be made of $\kappa_2 a(X) + b(X)$, note that to first order the extinction for $\Delta R$
%perturbations from a fixed $R^F_0$ is
%\begin{equation}
%A_X = A_V \left(a(X) + \left(\frac{1}{R^F_0} - \frac{\Delta R^F}{R^F_0} \right) b(X) \right).
%\end{equation}
%Then $\kappa^{-1}_2$ can this be associated with $\frac{1}{R^F_0} - \frac{\Delta R^F}{R^F_0}$.

@@

\color{orange}
As a check of the above interpretation, we generate ten simulated datasets with the same numbers and uncertainties as our data, based on a
SALT2 $x_1=0$, $c=0$ template and  \citetalias{1999PASP..111...63F} dust.  The supernovae
are assigned $E^F(B-V)$ parameters equal to the $B-V$ color excesses determined in our sample
(making an approximative correspondence between observed colors measured in our top-hat filters and the 
model parameter $E^F(B-V)$) , and $R^F$ values drawn from a log-normal
distribution with a mean of $\ln{2.5}$ and 0.05 standard deviation.  The $\alpha$, $\beta$, and $\eta$ parameters are set to zero, so that dust and
measurement uncertainty are the only sources of variation in magnitudes.  The input $R^F=2.5$ is centered
within the 68\% credible interval for $\kappa^{-1}_1$ of 2.38--2.69 from the results of the stacked 10 simulations.

@@

Figure~\ref{k0_med:fig} also shows the distribution of $\gamma^1_V k_1$, which can be attributed to 
extinction corrections due to deviations away from the canonical $\kappa_1$ value.   The distribution has a long
negative tail and is peaked at its positive edge, and by construction has a mean of zero.
\color{orange}
Note that the extreme outlier has large uncertainties, and is a manifestation of the unconstrained degree of freedom in this parameter set.
\color{black}
The implication is that while the bulk of supernovae have positive values associated with $R>1/\kappa_1$, there is a tail
of negative values associated with $R<1/\kappa_1$.  The plot of   $\gamma^0_V k_0$ versus $\gamma^1_V k_1$  shown in Figure~\ref{kk:fig} shows that
the positive edge in $\gamma^1_V k_1$ persists over the full range of  $\gamma^0_V k_0$
%\color{orange}
%While at low  $\gamma^0_V k_0$
%\color{black}
%On the other hand, lower values of $\gamma^1_V k_1$
%occur at larger values of  $\gamma^0_V k_0$.
These findings are consistent with previous results:
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F<3$ encountered by the SN~Ia population.
%\citet{2011ApJ...729...55F} and the above authors
%find that supernovae with large extinction 
%preferentially exhibit low $R^F$.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output11/gamma0_med.pdf}
   \includegraphics[width=2.8in]{output11/gamma1_med.pdf}
      \caption{
      \color{orange} Normalized ideogram and histogram 
      \color{black}
      of
median values of Left: $\gamma^0_V k_0$, which is associated with the $A^F_V$ of an $R^F \sim 1/\kappa_1$ dust model; Right: $\gamma^1_V k_1$,
which  is associated with extinction corrections due to deviations away from the canonical $\kappa_1$ value.
   \label{k0_med:fig}}
\end{figure}

@@

\subsection{SN 2014J in the Context of Our Model}
\label{sn2014j:sec}
SN~2014J   is one of several SNe~Ia that exhibit colors that imply a low $R_V<2.0$ \citep{2014ApJ...788L..21A, 2014MNRAS.443.2887F, 
2014arXiv1411.3332J,
2014ApJ...795L...4K, 2015ApJ...805...74B}.
The colors of this supernova can be studied in comparison to SN~2011fe, an object with similar
spectral evolution that 
suffers very low galactic and host reddening
\citep[this technique has been used in][]{2006MNRAS.369.1880E,2007AJ....133...58K,2008MNRAS.384..107E,2010AJ....139..120F, 2014ApJ...788L..21A,
2017arXiv170101422H}.
The overall higher photospheric velocities of
SN~2014J are unimportant  based on the findings of  \S\ref{results:sec}.

Data of SN~2014J are  analyzed using our model and using the parameter values found with the SNfactory sample.
The data for the color excesses  in $UBRi$  relative to $V$ at peak brightness  are taken from \citet{2014ApJ...788L..21A},
following their prescription of averaging measurements within 5-days of peak $B$ magnitude.
Their values 
$E(U_o-V_o) =   2.23 \pm   0.03$,
$E(B_o-V_o) =   1.28 \pm   0.04$,
$E(R_o-V)_o =  -0.47 \pm   0.03$,
$E(i_o-V_o) =  -0.92 \pm   0.03$
are plotted in Figure~\ref{sn2014j:fig}, where the subscript $o$ emphasizes that this is an observed color excess.
The central wavelengths of the filters are 3660, 4280, 5350, 6310, and 8100 \AA\ respectively, in contrast to the central
wavelengths 
3640, 4390, 5290, 6370, 7680 \AA\ of our synthetic bands.  Despite the imperfect correspondence our synthetic passbands
and those used for the observations,
these data are fit to our model
\begin{equation}
E(X_o-V_o) =  \left(\frac{\gamma^0_X}{\gamma^0_V}-1\right)e_0 +  \left(\frac{\gamma^1_X}{\gamma^1_V}-1\right)e_1,
\end{equation}
where we use the median values of the $\gamma$- and $\delta$-terms from Table~\ref{global:tab} and the fit
parameters $e_i $.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/sn2014j.pdf} 
   \caption{Top: Measured color excess for SN~2014J and the best-fit predictions from this article.   Bottom: Residual color excess
   between the data and model, with a line at zero shown for reference.
   This can be compared with the best-fit results of
   \citet{2014ApJ...788L..21A} shown in their Figure~3.
   \label{sn2014j:fig}}
\end{figure}

The best-fit parameters for SN~2014J are 
%----
$e_0= 2.60$, $ e_1=-1.89$ with covariance
%----
\begin{equation}
\begin{pmatrix}
\begin{array}{rr}
0.050 & 0.058 \\
0.058 & 0.315
\end{array}
\end{pmatrix}.
\end{equation}
The predicted values of observed $E(B_o-V_o)$ from the fit are shown in Figure~\ref{sn2014j:fig}, where they are found to
overlap with the data within the error bars.   
Our results can be compared with those of  \citet{2014ApJ...788L..21A} (Figure 3), who 
use
UV through NIR data to fit \citet{1999PASP..111...63F} dust with $R_V^F=1.4$ and $E(B-V)=1.37$.
Our model predictions provide a significantly better match to the data
despite the only approximate correspondence of the synthetic SNfactory and \citet{2014ApJ...788L..21A} observer passbands.

@@

\section{Supernova Model II: Two Extrinsic Parameters, One Intrinsic Parameter}
\label{model2:sec}
In \S\ref{model:sec} we show that the SNfactory data analyzed with an agnostic model with
two extrinsic parameters produce
magnitudes consistent
with two-parameter dust models that have been established for stars in the Milky Way, LMC, and SMC.
An extension to our model is now added to explore the possibility of
a possible additional
color parameter that could have its origins in supernova astrophysics.
The extended model referred to as Model~II is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_U+\alpha_U EW_{Ca} + \beta_U EW_{Si} + \eta_U \lambda_{Si} + \delta_U D\\
c_B+\alpha_B EW_{Ca} + \beta_B EW_{Si} + \eta_B \lambda_{Si} + \delta_B D \\
c_V+\alpha_V EW_{Ca} + \beta_V EW_{Si} + \eta_V \lambda_{Si} + \delta_V D\\
c_R+\alpha_R EW_{Ca} + \beta_R EW_{Si} + \eta_R \lambda_{Si} + \delta_R D\\
c_I+\alpha_I EW_{Ca} + \beta_I EW_{Si}+ \eta_I \lambda_{Si} + \delta_I D
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{U} k_0 +\gamma^1_{B} k_1 \\B +\gamma^0_{B} k_0 +\gamma^1_{B} k_1 \\
V+\gamma^0_{V} k_0+\gamma^1_{V} k_1\\R+\gamma^0_{R} k_0 + \gamma^1_{R} k_1\\I+\gamma^0_{I} k_0+\gamma^1_{I} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
In all aspects it is identical to the previous model except for the addition of the term $\delta D$,
which affects the mean
absolute magnitudes.  Thus
each supernova is given an intrinsic $D$ parameter that is linearly related to
absolute magnitudes through global coefficients $\delta$.  As with $k_0$ and $k_1$,  $D$ is a latent supernova parameter whose existence is inferred from
its effects on broadband magnitudes.

@@

To constrain the new degrees of freedom and degeneracies inherent in the model we impose additional constraints. We impose
\begin{equation}
\langle D\rangle=0,\,  \delta \cdot \gamma^0=0,\, \delta \cdot \gamma^1=0.
\end{equation}
The first of these constraints specifies the (arbitrary) definition of zero color.   The latter two confine possible magnitude
variations to be those that cannot be attributed to 
the $\gamma k$ terms (i.e.\ dust); this condition breaks the degeneracies between $\delta D$  
and $\gamma k$. 
Unlike for the $\gamma$'s, there is no condition on the sign of $\delta_U$,
so that the posterior has degenerate solutions with sign flips of $\delta$ and $D$.
We apply the condition $\delta_I>0$ to break this degeneracy,
and assume a flat prior for the three linear components of the subspace perpendicular to the $\gamma$ plane.
In the running of the Monte Carlo fit, the initial conditions for
$\gamma^0$ and $\gamma^1$ were set to be in the neighborhood of the best fit of \S\ref{model:sec}
to maintain their correspondence between the two models.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output23/ebv.pdf}
   \includegraphics[width=2.8in]{output23/ebv_delta.pdf}
      \caption{Left: Ideograms of the external $E(B-V)$ and
   internal $E_\delta(B-V) = (\delta_B-\delta_V)D$  contributions to color excess  for the supernovae in our sample.
   Right: Ideogram and histogram of the medians of $E_\delta(B-V)$ on an enlarged scale.
   \label{ebv:fig}}
\end{figure}
The significance of the $\delta$ parameters is shown pictorially in
Figure~\ref{deltacorner:fig}, which shows the $\delta$ posteriors.
Recall that fixing $\delta_X=0$ reverts this model into the two extrinsic-parameter model of \S\ref{model:sec}.
Assigning the Monte Carlo links to bins of width 0.5 in each $\delta$ parameter, the maximum number of links in a single bin is 303, the number of bins
with at least one element is 1987, and the bin that contains $\delta=0$ for all bands has zero elements; we thus claim a detection
of a third color-parameter at $> (1-5\times 10^{-4})$.
Figure~\ref{deltaratio:fig} shows  values of $\delta_X/\delta_I-1$ that  have significant non-zero values.
While the signal of the internal parameter is strong within the context of this model, ``model testing'' (comparing the
performance  between this model
and that of \S\ref{model:sec}), is challenging for Bayesian Monte Carlo analysis.  Evidence calculations are planned
for the next generation of STAN and the development of robust evidence algorithms is a subject of theoretical work.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
      \includegraphics[width=5.2in]{output23/deltaratio.pdf}
   \caption{Significance of the non-zero values of $\delta$ as demonstrated by the best-fit values of $\delta_X/\delta_I-1$ for $\delta$ for Model~II.
   A line at zero is shown for reference.
   \label{deltaratio:fig}}
\end{figure}

@@

\color{black}
The analysis is validated using simulated datasets with and without third color parameters: the input parameters are recovered within  uncertainties.
We present here the results of one of the validation tests.  The input values of the supernova parameters are the median values from the analysis of real data.
Ten realizations of simulated data are generated based on the data covariances, and then analyzed separately.   The posteriors of the ten realizations
are aggregated and shown in Figures~\ref{simglobal1:fig}--\ref{simglobal5:fig}.  The input parameters are recovered with negligible bias
relative to the statistical uncertainties.  There is a tendency for the slope parameters to be slightly underestimated, which is an expected consequence of 
the uncertainties in the predictor variable, which in turn lead to 
regression dilution.  Our analysis is susceptible to this bias because we consciously do not include distributions
of supernova parameters as part of our model.

@@

\section{Discussion}
